Point transect distance sampling in inlabru

### 2019-08-12 paper.tex
slight tweak

\citet{yuan_point_2017} presented a one-stage approach for analysing line-transect distance sampling within a Bayesian framework that simulateously estimates the detectability 
and spatial distribution of animals.  Key to their approach is the use of sparse Gaussian Markov random field (GMRF) approximations to continously indexed Gaussian random fields (GRF) and a log-Gaussian Cox point process model that avoids the need to bin data into counts and naturally sidesteps issues of zero-inflation and dispersion.  The observation process is formulated as a thinning of the log-Gaussian Cox process.  Both the log-intensity of points and the detection function were modelled using a stochastic partial differential equation approach \citep{lindgren_explicit_2011}.  The model was specified as a latent Gaussian model and fitted using integrated nested Laplace approximations using the \texttt{R-INLA} package \citep{rue_approximate_2009}.

Here we present a similar approach to \citet{yuan_point_2017} but instead focus on the case of point transect data.  Point transects require special treatment to account for the fact that the area surveyed increases with increasing distance from the observer.  Our case study is, to the best of our knowledge, the first analysis of point transect distance sampling data as a thinned point process.  However, the point process viewpoint is not new and has been applied numerous times to line transect data [[citations here]].

>>>>

\citet{yuan_point_2017} presented a one-stage approach for analysing line-transect distance sampling within a Bayesian framework that simulateously estimates the detectability 
and spatial distribution of animals.  Key to their approach is the use of sparse Gaussian Markov random field (GMRF) approximations to continously indexed Gaussian random fields (GRF) and a log-Gaussian Cox point process model that avoids the need to bin data into counts and naturally sidesteps issues of zero-inflation and dispersion.  The observation process is formulated as a thinning of the log-Gaussian Cox process.  Both the log-intensity of points and the detection function were modelled using a stochastic partial differential equation (SPDE) approach \citep{lindgren_explicit_2011}.  The model was specified as a latent Gaussian model and fitted using integrated nested Laplace approximations using the \texttt{R-INLA} package \citep{rue_approximate_2009}.

Here we present a similar approach to \citet{yuan_point_2017} but instead focus on the case of point transect data.  Point transects require special treatment to account for the fact that the area surveyed increases with increasing distance from the observer.  Our case study is, to the best of our knowledge, the first analysis of point transect distance sampling data formulated as a thinned point process.  However, the point process viewpoint is not new and has been applied numerous times to line transect data [[citations here]].  We also present the first use of sparse GMRF with point transect distance sampling data.

We demonstrate an application of a new iterative model fitting approach that allows the inclusion of model components that are non-linear in their parameters.  This allows us to model detectability with a parametric family of detection functions, an approach that more closely resembles standard distance sampling methods than the SPDE formulation presented in \citet{yuan_point_2017}.
